Supplemental Information: Continuous developmental changes in word recognition support language learning across early childhood

### 2025-04-04 paper/supplemental_information.Rmd
supplement update

To test for the differentiation of vocabulary growth based on initial reaction time, we used the package `nlme` to fit a logistic growth model to the production data. This model has two parameters for the logistic curve, a scale and an intercept. Both were allowed to interact with initial reaction time. We also included random effects of logistic intercept and scale by participant and a grouping term across datasets. This model showed a significant effect of initial reaction time on the intercept of the logistic growth curve, but not on its scale (see Table \ref{tab:nlme}).

<!-- % latex table generated in R 4.4.2 by xtable 1.8-4 package -->
<!-- % Thu Mar 13 13:10:30 2025 -->
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
\hline
& Value & Std Error & DF & $t$-value & $p$-value \\ 
\hline
Growth Intercept & 2.26 & 2.70 & 1766.00 & 0.84 & 0.40 \\ 
Growth Intercept $\times$ $t_0$ RT & 3.30 & 0.38 & 1766.00 & 8.59 & 0.00 \\ 
Growth Scale & 2.05 & 2.60 & 1766.00 & 0.79 & 0.43 \\ 
Growth Scale $\times$ $t_0$ RT & 0.28 & 0.38 & 1766.00 & 0.75 & 0.46 \\ 
\hline
\end{tabular}
\caption{Fixed effects estimates from logistic growth model. \label{tab:nlme}}
\end{table}

@@

# Latent Factor Growth

```{r}
fa3_model_long <- "
# measurement model
vocab_t1 =~ 1*prod_t1 + s1*comp_t1
accuracy_t1 =~ 1*acc_t1 + s2*acc_sd_t1
speed_t1 =~ 1*log_rt_t1 + s3*log_rt_sd_t1

vocab_t2 =~ 1*prod_t2 
accuracy_t2 =~ 1*acc_t2 + s2*acc_sd_t2
speed_t2 =~ 1*log_rt_t2 + s3*log_rt_sd_t2

vocab_t3 =~ 1*prod_t3 
accuracy_t3 =~ 1*acc_t3 + s2*acc_sd_t3
speed_t3 =~ 1*log_rt_t3 + s3*log_rt_sd_t3

vocab_t4 =~ 1*prod_t4 
accuracy_t4 =~ 1*acc_t4 + s2*acc_sd_t4
speed_t4 =~ 1*log_rt_t4 + s3*log_rt_sd_t4

# means for the latents
# vocab_t1 ~ 1
# vocab_t2 ~ 1
# vocab_t3 ~ 1
# vocab_t4 ~ 1
# accuracy_t1 ~ 1
# accuracy_t2 ~ 1
# accuracy_t3 ~ 1
# accuracy_t4 ~ 1
# speed_t1 ~ 1
# speed_t2 ~ 1
# speed_t3 ~ 1
# speed_t4 ~ 1

# autoregression for the latents
# vocab_t2 ~ 1 * vocab_t1
# vocab_t3 ~ 1 * vocab_t2
# vocab_t4 ~ 1 * vocab_t3
# accuracy_t2 ~ 1 * accuracy_t1
# accuracy_t3 ~ 1 * accuracy_t2
# accuracy_t4 ~ 1 * accuracy_t3
# speed_t2 ~ 1 * speed_t1
# speed_t3 ~ 1 * speed_t2
# speed_t4 ~ 1 * speed_t3

# regressions
accuracy_intercept =~ 1*accuracy_t1 + 1*accuracy_t2 + 1*accuracy_t3 + 1*accuracy_t4 
accuracy_slope =~ 1*accuracy_t1 + 2*accuracy_t2 + 3*accuracy_t3 + 4*accuracy_t4
speed_intercept =~ 1*speed_t1 + 1*speed_t2 + 1*speed_t3 + 1*speed_t4 
speed_slope =~ 1*speed_t1 + 2*speed_t2 + 3*speed_t3 + 4*speed_t4
vocab_intercept =~ 1*vocab_t1 + 1*vocab_t2 + 1*vocab_t3 + 1*vocab_t4 
vocab_slope =~ 1*vocab_t1 + 2*vocab_t2 + 3*vocab_t3 + 4*vocab_t4

# parameters for the regressions
vocab_intercept ~ 1
vocab_slope ~ 1
accuracy_intercept ~ 1
accuracy_slope ~ 1
speed_intercept ~ 1
speed_slope ~ 1

# residual variance for the latents
vocab_t1 ~~ NA*vocab_t1
vocab_t2 ~~ NA*vocab_t2
vocab_t3 ~~ NA*vocab_t3
vocab_t4 ~~ NA*vocab_t4
accuracy_t1 ~~ NA*accuracy_t1
accuracy_t2 ~~ NA*accuracy_t2
accuracy_t3 ~~ NA*accuracy_t3
accuracy_t4 ~~ NA*accuracy_t4
speed_t1 ~~ NA*speed_t1
speed_t2 ~~ NA*speed_t2
speed_t3 ~~ NA*speed_t3
speed_t4 ~~ NA*speed_t4

# residual variance for the other latents
vocab_intercept ~~ NA*vocab_intercept
vocab_slope ~~ NA*vocab_slope
accuracy_intercept ~~ NA*accuracy_intercept
accuracy_slope ~~ NA*accuracy_slope
speed_intercept ~~ NA*speed_intercept
speed_slope ~~ NA*speed_slope

# set observed intercepts to zero - these should go into the latents
prod_t1 ~ 0
prod_t2 ~ 0
prod_t3 ~ 0
prod_t4 ~ 0
comp_t1 ~ 0 
log_rt_t1 ~ 0
log_rt_t2 ~ 0
log_rt_t3 ~ 0
log_rt_t4 ~ 0
log_rt_sd_t1 ~ 0 
log_rt_sd_t2 ~ 0
log_rt_sd_t3 ~ 0 
log_rt_sd_t4 ~ 0 
acc_t1 ~ 0 
acc_t2 ~ 0 
acc_t3 ~ 0 
acc_t4 ~ 0 
acc_sd_t1 ~ 0
acc_sd_t2 ~ 0
acc_sd_t3 ~ 0
acc_sd_t4 ~ 0
"

fit3_long <- sem(fa3_model_long, d_sub_wide_fine, std.lv=TRUE, missing='fiml')

summary(fit3_long, fit.measures=TRUE, standardize=TRUE)

```{r}
layout <- read.csv(here("misc/layout_dt3.csv"), header = FALSE)
graph_sem(model = fit3_long, text_size = 3, layout=t(layout))

<!-- # Short Window Analysis -->

<!-- Need to redo all figures using short windows ugh.  -->
# References

>>>>

To test for the differentiation of vocabulary growth based on initial reaction time, we used the package `nlme` to fit a logistic growth model to the production data. This model has two parameters for the logistic curve, a scale and an intercept. Both were allowed to interact with initial reaction time. We also included random effects of logistic intercept and scale by participant and a grouping term across datasets. This model showed a significant effect of initial reaction time on the intercept of the logistic growth curve, but not on its scale (see Table \ref{tab:nlme}). 
```{r}
# xtable(summary(mod_nlme)$tTable)
```

<!-- % latex table generated in R 4.4.2 by xtable 1.8-4 package -->
<!-- % Thu Mar 13 13:10:30 2025 -->
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
\hline
& Value & Std Error & DF & $t$-value & $p$-value \\ 
\hline
Growth Intercept & 2.26 & 2.70 & 1766.00 & 0.84 & 0.40 \\ 
Growth Intercept $\times$ $t_0$ RT & 3.30 & 0.38 & 1766.00 & 8.59 & < 0.0001 \\ 
Growth Scale & 2.05 & 2.60 & 1766.00 & 0.79 & 0.43 \\ 
Growth Scale $\times$ $t_0$ RT & 0.28 & 0.38 & 1766.00 & 0.75 & 0.46 \\ 
\hline
\end{tabular}
\caption{Fixed effects estimates from logistic growth model. \label{tab:nlme}}
\end{table}

@@

# References
